


Just getting graphs for the XCI-skew project
```{r}
library(VGAM) 
library("gridExtra")
library(dplyr)
library(ggplot2)
library(scales)

```


load the max-powered SNPs for each sample, and the statistics dataframe

```{r}
load("../data/all_v8_GTEx_gene_filtered.skew.est.max.genes.Rdata")
load("../data/v8_GTEx_skew_and_stats_df.Rdata")




```



```{r}
folded <- function(x) { apply( cbind(x,1-x), 1, max)} 

```

Making basic binomial plots for the figure 1 graphic. Demonstrating the binomial distribution of XCI

####################################
Figure 1 graphs in panels a and b
###################################


```{r}
#Binomial from 16 cells
par(pty = 's')
p = .5
n = 16
x = 0:n
test = dbinom(x, size = n, prob = p)
plot(x/n, test, xlab = 'Xm / Xm + Xp', ylab = 'probability', type ='b', lwd=5, pch=19, cex.lab = 1.5, cex.axis=1.25)
box(lwd = 3)



#Binomial from 32 cells
n = 32
x = 0:n
test = dbinom(x, size = n, prob = p)
plot(x/n, test, xlab = 'Xm / Xm + Xp', ylab = 'probability', type ='b', lwd=5, pch=19, cex.lab = 1.5, cex.axis = 1.25)
box(lwd = 3)


#Binomial from 8 cells
n = 8
x = 0:n
test = dbinom(x, size = n, prob = p)
plot(x/n, test, xlab = 'Xm / Xm + Xp', ylab = 'probability', type ='b', lwd=5, pch=19, cex.lab = 1.5, cex.axis = 1.25)
box(lwd = 3)

```


Simulate data for a 75% skewed sample, unfolded and folded for the example in Figure 2. Add the folded normal model to it as well

#########################
Figure 2 panel a
#########################

```{r}
mat_skew = .75
pat_skew = 1 - mat_skew
sd_skew = .05

num_genes = 10000

mat_exp = rnorm(n = num_genes, mean = mat_skew, sd = sd_skew)
pat_exp = rnorm(n = num_genes, mean = pat_skew, sd = sd_skew)

hist(mat_exp, xlim = c(0,1), breaks = seq(0,1,.025), freq = F)
hist(pat_exp, xlim = c(0,1), breaks = seq(0,1,.025), freq = F, add = T)

folded_skews = c(mat_exp, pat_exp)
folded_skews = folded(folded_skews)

hist(folded_skews, xlim = c(.5,1))



exp_skew_df = data.frame(skews = c(c(mat_exp, pat_exp), folded_skews), label = c(rep('Unfolded', length = num_genes*2), rep('Folded', length = num_genes*2)))



gg_2 <- ggplot(exp_skew_df[exp_skew_df$label == 'Unfolded', ], aes(x=skews)) + geom_histogram(binwidth = .05, colour="black", fill = "#06e3ff", aes(y=..density..)) + 
  ggtitle('Unfolded skews') +
  scale_x_continuous(limits = c(-0.01, 1.01)) + 
  theme( plot.title=element_text(hjust=.5, size=15), axis.text.y = element_text(size=15), axis.title.y = element_text(size=15),
         axis.text.x = element_text(size=15), axis.title.x = element_text(size=15), 
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.border = element_rect(colour = "black", fill=NA, size=1), panel.background = element_rect(fill = 'white')) +
  xlab('Reference skew')

gg_2 = gg_2 + geom_vline(xintercept = .5, linetype="dashed", color = "red", size=1)
gg_2


gg_2f <- ggplot(exp_skew_df[exp_skew_df$label == 'Folded', ], aes(x=skews)) + geom_histogram(binwidth = .025, colour="black", fill = 'white', aes(y=..density..)) +
  ggtitle('Folded skews') + scale_x_continuous(limits = c(.499, 1.01))+ 
  theme( plot.title=element_text(hjust=.5, size=15), axis.text.y = element_text(size=15), axis.title.y = element_text(size=15),
         axis.text.x = element_text(size=15), axis.title.x = element_text(size=15), 
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.border = element_rect(colour = "black", fill=NA, size=1), panel.background = element_rect(fill = 'white')) + 
  xlab('Folded reference skew')

gg_2f <- gg_2f + stat_function(fun=dfoldnorm, color="red", args=list(mean=mat_skew, sd=sd_skew), size = 1) +
  geom_vline(xintercept = .5, linetype="dashed", color = "red", size=1)
gg_2f

grid.arrange(gg_2, gg_2f, nrow=1)
skewed_gg = arrangeGrob(gg_2, gg_2f, nrow=1)

```

Plot unfolded and folded examples of real tissue gene skews for a range of skewed tissues

```{r}
example_skews = c(.55,.65,.75,.85)
limit = .04
i = 4
potential_samples = skew_and_stats_df$sample_index[(skew_and_stats_df$skew >= example_skews[i] - limit) & (skew_and_stats_df$skew <= example_skews[i] + limit) & skew_and_stats_df$num_snps >= 50]
length(potential_samples)

for(j in 1:length(potential_samples)){
  sample_data = list.skew.max[[potential_samples[j]]]
  sample_ref_skews = sample_data$C.1
  est_skew = skew_and_stats_df$skew[skew_and_stats_df$sample_index ==potential_samples[j] ]
  hist(sample_ref_skews, breaks = seq(0,1,.05), xlim = c(0,1), main = sprintf('%i: %1.3f',potential_samples[j], est_skew) )
}
```

```{r}

example_sample_ids = c(4003, 602, 4647, 1368)
example_sample_skews = skew_and_stats_df$skew[match(example_sample_ids,skew_and_stats_df$sample_index)]
example_sample_skews_sd = skew_and_stats_df$skew_sigma[match(example_sample_ids,skew_and_stats_df$sample_index)]

skews = c()
folded_label = c()
sample_label = c()
for(i in 1:length(example_sample_ids)){
  #Get the reference skews
  sample_data = list.skew.max[[example_sample_ids[i]]]
  sample_ref_skews = sample_data$C.1
  skews_add = c(sample_ref_skews, folded(sample_ref_skews))
  skews = c(skews, skews_add)
  folded_label_add = c(rep('Unfolded', length = length(sample_ref_skews)), rep('Folded', length = length(folded(sample_ref_skews))) )
  folded_label = c(folded_label, folded_label_add)
  text = as.character(sprintf('Sample skew: %1.3f', example_sample_skews[i]))
  sample_label = c(sample_label, rep(text, length=length(skews_add)))
}

multi_skew_tissue_df = data.frame(skews = skews, folded_label = folded_label, sample_label = sample_label)
multi_skew_tissue_df


```

```{r}
sample_labels = unique(multi_skew_tissue_df$sample_label)



make_unfolded_ggplot = function(sample_label){

  data_index = multi_skew_tissue_df$folded_label == 'Unfolded' & multi_skew_tissue_df$sample_label == sample_label
  g = ggplot(multi_skew_tissue_df[data_index, ], aes(x = skews)) + geom_histogram(binwidth = .05, colour="black", fill = "#06e3ff") +
    geom_vline(xintercept = .5, linetype="dashed", color = "red", size=1) +
    ggtitle(sample_label) + 
    scale_x_continuous(limits = c(-.01, 1.01))+ 
    theme( plot.title=element_text(hjust=.5, size=15), axis.text.y = element_text(size=15), axis.title.y = element_text(size=15),
           axis.text.x = element_text(size=15), axis.title.x = element_text(size=15), 
           panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
           panel.border = element_rect(colour = "black", fill=NA, size=1), panel.background = element_rect(fill = 'white')) + 
    xlab('Reference skew')
  return(g)
}

g1 = make_unfolded_ggplot(sample_labels[1]) 
g2 = make_unfolded_ggplot(sample_labels[2]) 
g3 = make_unfolded_ggplot(sample_labels[3])  
g4 = make_unfolded_ggplot(sample_labels[4]) 

grid.arrange(g1, g2,g3,g4, nrow=1)
skewed_gg = arrangeGrob(g1, g2,g3,g4, nrow=1)




make_folded_ggplot = function(sample_label, mean_skew, sd_skew){

  data_index = multi_skew_tissue_df$folded_label == 'Folded' & multi_skew_tissue_df$sample_label == sample_label
  g = ggplot(multi_skew_tissue_df[data_index, ], aes(x = skews)) + geom_histogram(binwidth = .025, colour="black", fill = 'white', aes(y=..density..)) +
    geom_vline(xintercept = .5, linetype="dashed", color = "red", size=1) +
    ggtitle(sample_label) + 
    scale_x_continuous(limits = c(.499, 1.01))+ 
    stat_function(fun=dfoldnorm, color="red", args=list(mean=mean_skew, sd=sd_skew), size = 1) +
    theme( plot.title=element_text(hjust=.5, size=15), axis.text.y = element_text(size=15), axis.title.y = element_text(size=15),
           axis.text.x = element_text(size=15), axis.title.x = element_text(size=15), 
           panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
           panel.border = element_rect(colour = "black", fill=NA, size=1), panel.background = element_rect(fill = 'white')) + 
    xlab('Folded reference skew')
  return(g)
}


g1_f = make_folded_ggplot(sample_labels[1], example_sample_skews[1], example_sample_skews_sd[1] ) 
g2_f = make_folded_ggplot(sample_labels[2], example_sample_skews[2], example_sample_skews_sd[2] ) 
g3_f = make_folded_ggplot(sample_labels[3], example_sample_skews[3], example_sample_skews_sd[3] )  
g4_f = make_folded_ggplot(sample_labels[4], example_sample_skews[4], example_sample_skews_sd[4] ) 

grid.arrange(g1_f, g2_f,g3_f,g4_f, nrow=1)
skewed_gg = arrangeGrob(g1_f, g2_f,g3_f,g4_f, nrow=1)

```

###########################
Figure 2 panel b
###########################


```{r}
grid.arrange(g1, g2,g3,g4, g1_f, g2_f,g3_f,g4_f, nrow=2)
combined_skewed_gg = arrangeGrob(g1, g2,g3,g4, g1_f, g2_f,g3_f,g4_f, nrow=2)

```


Plot different binomial distributions centered at .75 for the tissue specification graphic

########################
Figure 5 panel c
########################


```{r}

skew = .75
n_1 = 16
n_2 = 64
par(pty = 's')

skews_1 = 0:n_1
binom_1 = dbinom(skews_1, size = n_1, prob = skew )
plot(skews_1/n_1, binom_1, type ='b', lwd=5, pch=19, cex.lab = 1.5, cex.axis = 1.25, xlim = c(.5, 1))
abline(v = .75, col = 'red')


par(pty = 's')

skews_2 = 0:n_2
binom_2 = dbinom(skews_2, size = n_2, prob = skew )
plot(skews_2/n_2, binom_2, type ='b', lwd=5, pch=19, cex.lab = 1.5, cex.axis = 1.25, xlim = c(.5, 1))
abline(v = .75, col = 'red')

```


```{r}

head(skew_and_stats_df)
dim(skew_and_stats_df)
```



























